---
title: "p8105_hw3_yc4016"
output: github_document
date: "2022-10-07"
author: "Yishan Chen"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 1

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Description
* The size of the dataset `instacart` is `r dim(instacart)` 
* The dataset `instacart` contains `r nrow(instacart)` rows and `r ncol(instacart)` with each row resprenting a single product from an instacart order.
* The key variables include `r names(instacart)`
* In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.


#### Question Prompt
How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

#### Description
* In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.


#### Question Prompt
Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


#### Question Prompt
Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```


#### Question Prompt
Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```


# Problem 2

#### Question Prompt
Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accelerometers = 
  read_csv("data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_each_minute",
    names_prefix = "activity_",
    values_to = "counts"
  ) %>% 
  mutate(
    weekday_vs_weekend = if_else(day %in% c("Saturday", "Sunday"), 1, 0),
    weekday_vs_weekend = as.factor(weekday_vs_weekend)
         ) 
  
accelerometers
```

#### Description
* The size of the dataset `accelerometers` is `r dim(accelerometers)` 
* The dataset `accelerometers` contains `r nrow(accelerometers)` rows and `r ncol(accelerometers)` columns
* The dataset `accelerometers` contains `r nrow(accelerometers)` observations
* The key variables include `r names(accelerometers)`


#### Question Prompt
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?


```{r}
accelerometers_table =
accelerometers %>% 
  group_by(week, day) %>% 
  summarize(total_activity = sum(counts)) %>% 
  pivot_wider(
    names_from = "day",
    values_from = "total_activity") %>% 
  select (Monday, Tuesday, Wednesday, Thursday, Friday, Saturday, Sunday) %>% 
  knitr::kable()

accelerometers_table
```

#### Description
* It seems that the weekends has lower value than weekdays. 
* At Saturday of week 4 and 5 the number is pretty small.


#### Question Prompt 
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accelerometers_plot = 
accelerometers %>% 
  ggplot(aes(x = activity_each_minute, y = counts, color = day)) +
  geom_line(alpha = .5) +
  geom_smooth(se = FALSE) +
  labs(title = "The 24-hour activity time courses for each day", 
       x = "activity time", 
       y = "activity value")
  scale_x_continuous(
    breaks = c(0, 6, 12, 18, 24), 
    labels = c("00:00","06:00","12:00","18:00","24:00"))

accelerometers_plot
```



